\PYGEOMETRY{}: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-09-09 public/paper/pyg4ometry.tex
Fix error in the overlap checking algorithm section.
The for loop was inside the branch on whether or not the daughter overlaps
with the mother.  Suggesting that the other daughters were only checked if the mother
overlapped with the given daughter.
paper update

\subsection{Materials}
\PYGEOMETRY{} converts GDML/Geant4 materials to FLUKA \verb|MATERIAL| and \verb|COMPOUND| cards.
Geant4 has a class \cpinline{G4Material} to assign material state (density, physical state, temperature and pressure) 
to a logical volume. \cpinline{G4Material} has two main constructors, the first were a atomic number is supplied and 
the second is when \cpinline{G4Element} instances and relative atomic or mass abundances are provided. The {\em simple}
\cpinline{G4Material} is converted to a \fluka{material} card, whilst the {\em element} \cpinline{G4Material}  is converted to a 
\fluka{compound} card. A similar issues existing when converting \cpinline{G4Element} to FLUKA, as \cpinline{G4Element}
can either be simple, so defined only by atomic number and mass or composite and defined by a admixture defined by 
relative abundances of \cpinline{G4Isotope}. A similar mapping is performed so if an \cpinline{G4Element} is simple it is 
directly converted to a FLUKA \fluka{MATERIAL} card, and the {\it isotope} \cpinline{G4Element} is converted to a \fluka{compund}
card.

@@

\section{STL and CAD to GDML conversion}
STL and CAD conversion are closely related. In both cases the solid 
(in the case of STL) and solids (in the case of CAD) are converted to 
tessellated solid(s). STL is a relatively simple file format that can
be loaded using pure Python. As STL files typically only contain a single 
solid the \pyinline{stl.Reader} provides a single solid and not a logical volume
as with the other file readers.

@@

An example geometry representing a dipole electromagnet, consisting of three parts was created in Autodesk Fusion 360
and saved as a STEP file, \PYGEOMETRY{} produces the output in Figure~\ref{fig:cad-gdml}
\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/SectorBend.pdf}
\caption{Example conversion of a simple CAD (STEP) geometry (sector bend dipole electromagnet) to GDML, left panel:  
the original model in FreeCAD, right panel: the GDML geometry 
viewed in \PYGEOMETRY{}. }
\label{fig:cad-gdml}
\end{center}
\end{figure*}

\begin{enumerate}
\item Mesh overlap and shrinking algorithm 
\item Surface conversion and decomposition algorithm (future release)
\end{enumerate}

\section{Complete simulation example}
\PYGEOMETRY{} is designed to be as flexible as possible and offer the user a wide range of usage styles, input files 
and workflows. A fictitious beam line was created to demonstrate the capabilities of \PYGEOMETRY, this creates a composite 
{\it scene} which consists of the different examples presented in this paper. The beam line consists of a vacuum chamber 
(modelled in \PYGEOMETRY{}), a vacuum gate value (STL from the manufacturer), a triplet of quadrupole magnets (exported 
from BDSIM), a sector bend dipole electromagnet (created in Autodesk Fusion 360) and finally a Faraday cup (modelled in 
Flair). Each different file is loaded using \PYGEOMETRY{} reader class and the placed as a physical volume. The final composite 
geometry is shown in Figure~\ref{fig:model}
%
\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/model.pdf}
\caption{Complete compositing example using \PYGEOMETRY{}.}
\label{fig:model}
\end{center}
\end{figure*}

Having geometry wrapped in a suitable API allows a wide range of processes to be performed simply and potentially 
programmatically if it must be performed for large numbers of objects. The processes include; 
\begin{enumerate}
\item merging registries, 
\item removing volumes (de-featuring),
\item editing solid parameters,
\item replacing logical volume materials 
\item converting logical volumes to assembly volumes.
\end{enumerate}
The merging of registries and removing volumes is required to create the example shown in Figure~\ref{fig:model}. 
Each sub-component is stored in a separate registry and these have to combined without any GDML tags clashing. 
The FLUKA to GDML conversion creates logical volumes which are not generally required in Geant4, so for example
the air surrounding the Faraday cup, which is needed to specify a FLUKA geometry is converted to GDML but can be 
safely removed. Examples of other workflow can be found in the \PYGEOMETRY{} online manual
.
\section{Conclusions and discussion}
The authors believe that tools to quickly create, either ab-initio or by conversion geometry usable in 
particle transport Monte Carlo programmes will save significant amounts of user effort and ultimately 
yield more accurate simulations. \PYGEOMETRY{} is a relatively complete implementation of a geometry 
creation tool, whilst heavily internally based on Geant4 and GDML can have utility for users of all RT 
codes. \PYGEOMETRY{} can clearly be extended to other formats or applications. As for applications, examples 
include  the  generation triangular mesh structures for GPU accelerated photon tracking in liquid noble dark 
matter detectors and geometries for event visualisation. Although not implemented in the current version, 
a C++ output writer can be quickly implemented to generate geometry which can be compiled into a Geant4 
application.

\begin{enumerate}
\item Graphical front end viewer
\item C++ output
\item CAD BREP to FLUKA converter (a la McCAD) 
\item Integration of other tools (DAGMC) 
\item Far from perfect conversions
\item Export mesh for other applications (GPU optics)
\item Convergence FLUKA/Geant4 - Moira 
\item Replace G4 polyhedron CSG
\item Not for large scale conversion
\item Export for other RTMC codes (MCNP, etc) 
\item Performance in Geant4/FLUKA
\end{enumerate}

>>>>

\subsection{Materials}
\PYGEOMETRY{} converts GDML/Geant4 materials to FLUKA \verb|MATERIAL| and \verb|COMPOUND| cards.
Geant4 has a class \cpinline{G4Material} to assign material state (density, physical state, temperature and pressure) 
to a logical volume. \cpinline{G4Material} has two main constructors, the first were a atomic number is supplied and 
the second is when \cpinline{G4Element} instances and relative atomic or mass abundances are provided. The {\em simple}
\cpinline{G4Material} is converted to a \fluka{material} card, whilst the {\em element} \cpinline{G4Material}  is converted to a 
\fluka{compound} card. A similar issues existing when converting \cpinline{G4Element} to FLUKA, as \cpinline{G4Element}
can either be simple, so defined only by atomic number and mass or composite and defined by a admixture defined by 
relative abundances of \cpinline{G4Isotope}. A similar mapping is performed so if an \cpinline{G4Element} is simple it is 
directly converted to a FLUKA \fluka{MATERIAL} card, and the {\it isotope} \cpinline{G4Element} is converted to a \fluka{compund}
card. Geant4 also defines a set of standard materials \cite{NIST} from NIST so a user can just specify something like 
\pyinline{G4_STAINLESS-STEEL}, \PYGEOMETRY contains a matching database and creates the appropriate FLUKA cards during 
the conversion.

@@

\section{STL and CAD to GDML conversion}
STL and CAD conversion are closely related. In both cases the solid 
(in the case of STL) and solids (in the case of CAD) are converted to 
tessellated solid(s). STL is a relatively simple file format that can
be loaded using pure Python. As STL files typically only contain a single 
solid the \pyinline{stl.Reader} provides a single solid and not a logical volume
as with the other file readers.
%
\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/SectorBend.pdf}
\caption{Example conversion of a simple CAD (STEP) geometry (sector bend dipole electromagnet) to GDML, left panel:  
the original model in FreeCAD, right panel: the GDML geometry 
viewed in \PYGEOMETRY{}. }
\label{fig:cad-gdml}
\end{center}
\end{figure*}

@@

An example geometry representing a dipole electromagnet, consisting of three parts was created in Autodesk Fusion 360
and saved as a STEP file, \PYGEOMETRY{} produces the output in Figure~\ref{fig:cad-gdml}
\begin{enumerate}
\item Mesh overlap and shrinking algorithm 
\item Surface conversion and decomposition algorithm (future release)
\end{enumerate}

\section{Complete simulation example}
\PYGEOMETRY{} is designed to be as flexible as possible and offer the user a wide range of usage styles, input files 
and workflows. A fictitious beam line was created to demonstrate the capabilities of \PYGEOMETRY, this creates a composite 
{\it scene} which consists of the different examples presented in this paper. The beam line consists of a vacuum chamber 
(modelled in \PYGEOMETRY{}), a vacuum gate value (STL from the manufacturer), a triplet of quadrupole magnets (exported 
from BDSIM), a sector bend dipole electromagnet (created in Autodesk Fusion 360) and finally a Faraday cup (modelled in 
Flair). Each different file is loaded using \PYGEOMETRY{} reader class and the placed as a physical volume. The final composite 
geometry is shown in Figure~\ref{fig:model}. It must be noted when this complete geometry is written to GDML and loaded into 
Geant4 it cannot be visualised in anything but the ray tracer because of limitations in the OpenGL visualisation in Geant4. Another 
important note is that this geometry cannot be converted to FLUKA as it contains tessellated solids (both the STL gate valve and
dipole magnet). 
%
\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/model.pdf}
\caption{Complete compositing example using \PYGEOMETRY{}.}
\label{fig:model}
\end{center}
\end{figure*}

Having geometry wrapped in a suitable API allows a wide range of processes to be performed simply and potentially 
programmatically if it must be performed for large numbers of objects. The processes include; 
\begin{enumerate}
\item merging registries, 
\item removing volumes (de-featuring),
\item editing solid parameters,
\item replacing logical volume materials 
\item converting logical volumes to assembly volumes.
\end{enumerate}
The merging of registries and removing volumes is required to create the example shown in Figure~\ref{fig:model}. 
Each sub-component is stored in a separate registry and these have to combined without any GDML tags clashing. 
The FLUKA to GDML conversion creates logical volumes which are not generally required in Geant4, so for example
the air surrounding the Faraday cup, which is needed to specify a FLUKA geometry is converted to GDML but can be 
safely removed. Examples of other workflows or geometry manipulation processes can be found in 
the \PYGEOMETRY{} online manual.

\section{Conclusions and discussion}
The authors believe that tools to quickly create, either ab-initio or by conversion geometry usable in 
particle transport Monte Carlo programmes will save significant amounts of user effort and ultimately 
yield more accurate simulations. \PYGEOMETRY{} is a relatively complete implementation of a geometry 
creation tool, whilst heavily internally based on Geant4 and GDML can have utility for users of all RT 
codes. \PYGEOMETRY{} can clearly be extended to other formats or applications. As for applications, examples 
include  the  generation triangular mesh structures for GPU accelerated photon tracking in liquid noble dark 
matter detectors and geometries for event visualisation. Although not implemented in the current version, 
a C++ output writer can be quickly implemented to generate geometry which can be compiled into a Geant4 
application.

\begin{enumerate}
\item Coherent, uniform interface to existing tools
\item Graphical front end viewer
\item C++ output
\item CAD BREP to FLUKA converter (a la McCAD) 
\item Integration of other tools (DAGMC) 
\item Far from perfect conversions
\item Export mesh for other applications (GPU optics)
\item Convergence FLUKA/Geant4 - Moira 
\item Replace G4 polyhedron CSG
\item Not for large scale conversion
\item Export for other RTMC codes (MCNP, etc) 
\item Performance in Geant4/FLUKA
\item AR/VR export (to OBJ or XXX)
\item Where to put manual, unit tests, installation, git? 
\item range of current users (RAL, CERN, LUXE, QUB, IHEP-CEPC, etc) 
\item Paraview loader 
\item 2D section viewer 
\item PHITS, MCNP(X)
\end{enumerate}

\section{Acknowledgements}
This work has been supported by STFC grants XXX and XXX. Multiple undergraduate and masters students 
have contributed to the code for their degree project work, Simon Williams, Benjamin Shellswell, Joshua Albrecht. 
STB has benefited from discussions with and insight of Vasilis Vlachoudis (CERN) on FLUKA geometry and FLAIR. 
Cédric Hernalsteens (ULB/CERN) and Alistair Butcher (RHUL) also made useful contributions to the code.
